Oriented Gradient Histogram applied to P300 Detection

### 2017-12-10 response.letter.tex

A couple of finale observations. First, it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference. Another issue has to do with the way the manuscript is written, sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5), the manuscript surely benefit from a careful review before submitting.

\subsection*{\ovalbox{General Comments}}
I have read the manuscript "EEG characterization and classification based on Histogram of Gradients of Signal Plots" with great interest since its title promised a solution to an important problem. In spite of presenting an interesting methodology and some interesting issues related to EEG characterization, I consider the manuscript is not ready for publication. 
\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your feedback. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.1) the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. The title of the article was changed accordingly to your suggestion.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

2.2) there is evidence of the benefits of working with these signals P300, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
You're absolutely right, thank you very much for your comment.  We published in Ramele et al 2016 the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to decode Motor Imagery.  We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns. We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.

% Como haces para que las referencias que estan aca en la respuesta apunten al articulo ???
%An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex
\subsection*{\ovalbox{Third Paragraph}}

@@

3.1) Another issue with the manuscript is its mathematical vagueness. For example, the variable T is used as a vector or two-tuple but its notation is different from the notation the authors use for vectors

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  We worked hard to improve greatly all the mathematical description of the method. Section FEATURE was modified completely as well as section PARAMETERS.  Section SIGNAL PLOT was revised and its mathematical terms corrected.}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.2) Moreover, there is something wrong in equation (4)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We appreciate a lot your truthful comment.  Equation (4) is very important for us because it is the essence of the method.  We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.  We reviewed the equation, clarified their terms and hope it is now more clear.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.3) and something inconsistent between equations (7) and (8)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We have revised the manuscript and corrected equations (7) and (8), clarifying more precisely that they are part of the proposed method and that they help to understand how signal properties are related to the SIFT's geometric parameters.
We thank you for noticing this, that led to a better clarification of mathematical terminology.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.4) Also, the authors go back and forth the digital space and the continuous space and although there is an equivalence, the manuscript could benefit from a clearer explanation.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  We modified the equation to emphasize the point that is actually calculated in digital space. We also clarified that image gradients are also calculated based on finite differences.
With this we hope we have addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

4.1) Another issue that I find worrisome is the design of experiments.\\

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. 
We fully addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

4.2) it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your suggestion. The original dataset was published at the BNCI-Horizon website, which is very popular in BCI research.  Our approach emphasize the structure of the signal in terms of their morphological components in their visual description. THIS IS TOUGH TO JUSTIFY 
In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to understand the method.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

4.3) Also, the authors speculate that their results are not that good because they did not take into account the latency and magnitude of the signals, there is previous evidence in the literature that the P300 suffers from these, and other, problems and, thus, one wonders why the authors did not anticipate such problems.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
THIS IS EVEN TOUGHER.

ACA LE PUEDO PONER QUE HAY POCOS METODOS QUE PERMITAN ENGANCHAR BIEN QUE EL METODO

All these points were modified according to the suggestions.
We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

\subsection*{\ovalbox{Fifth Paragraph}}

5.1) it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
The Results section was modified to clarify more precisely the expectation about the EEG channels where a visual P300 detection task would be obtained.  We added references and improved the section to clarify this point more thoughtfully.
We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

5.2)  issue has to do with the way the manuscript is written

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you a lot for this important comment.  We re-read all the manuscript and asked to collegues to proofread the article.  They gave us interesting suggestions that helped us to improve different sections.  We would be very pleased if you find it now easier to read it.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

5.3) sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you very much for this information.  We reviewed completely that equation (it is now equation (6) and equation (7)) and we included a step-by-step procedure that describes the algorithm that we propose and we used.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

### 2017-12-10 article.tex
Reescribi la sección del nuevo dataset, revisé la sección de los parámetros, cambie algunas imagenes, incluí algo de informacion de la selección del canal.  Complete parte de la carta de respuesta para el revisor. Pequeñas correcciones aqui y alla.

This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained. Section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.  Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.  Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Near Neighbor  and the final Section shows results and discussion where we expose our remarks, conclusions and future work.

@@

The P300~\citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.  It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and it is consistent across different subjects. It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\citep{Hu2010}.  This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}. Fig.~\ref{fig:p300matrix} shows an example of a Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.   Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.

@@

The first step consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
The processing pipeline starts by applying a notch filter to the raw signal, a 
$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.

@@

The aim is to identify the selected letter from the matrix. Previously, during training phase, descriptors from labeled segments are extracted.  These descriptors are the P300 templates which are grouped in a template set called $ T $.  Rows are labeled 1-6, whereas columns are labeled 7-12.  The process has the following steps:

\begin{enumerate}
\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively. The parameter $k_a$ is an input of the algorithm.
\item Compute  
\begin{equation}
x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
\label{averaging}
\end{equation}

@@

\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.

\item Match with the Template $T$ by computing

@@

where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.  By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $. The  $k$ value is an input parameter of the algorithm.

@@

Figure~\ref{fig:classification} shows a schema of this process. 
\begin{figure}
\centering
\includegraphics[width=15cm]{classificationgraph.pdf}
\caption{Single trial segments are averaged for the 6 rows and 6 columns.}
\label{fig:classification}
\end{figure}

@@

To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, is used. An own dataset with  the same experimental conditions is generated. Both of them are utilized to perform an offline BCI Simulation to decode the spelled words from the provided signals.

The algorithm was implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).

@@

\subsubsection{P300 ALS Public Dataset}

@@

The recorded dataset was sampled at 256 Hz and it consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.

@@

@@

Parameters are selected according to the experimental protocol. It is define in Section~\ref{Feature}, the patch size is $S_P = 3s.4 \times 3s.4$ .
As the P300 event latency and amplitude vary greatly between subjects, it is necessary to utilize a size patch $S_P$ that be able to capture an entire transient event. With this purpose the $s$ value election is essential.
%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
We proposed the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} for compute the scale value in horizontal and vertical directions, respectively. 
\begin{equation}
s_x = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
\label{eq:mapping2}
\end{equation}

@@

\noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch. In this way the whole event is capture. But we need a squared patch, then we discovered that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, span of $ \lambda = 0.56$ seconds.  The resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs}\gamma$ in horizontal direction. Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).

% esta figura no está comentada en el texto.
\begin{figure}[H]
\centering
\includegraphics[width=10cm]{localpatch.png}
\caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
\label{fig:sampledescriptor2}
\end{figure}

We replicat the experiment on healthy subjects using g.Tec g.Nautilus 8 channels.  The experiment protocol is  the same and we 20 subjects, age $ 30$, 5 females and 15 males, 6 right handed and the remaining left handed.

Participants were recruited voluntarily and the experiment were conducted under anonymous information in accordance with the declaration of Helsinki.  No monetary compensation was delivered and all participants agreed and signed a written informed consent.  All healthy participants had normal or corrected-to normal vision and no history of neuropsychiatric disorders. The experiment was performed with 13 subjects, 7 males, 6 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.

EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems. The reference positions were placed at the right ear. The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed). Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz). Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz. The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system. Table~\ref{tab:results} summarizes the acquisition settings for each system.

@@

Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.  From the 7 trials for each subject, the first 3 were used as training, and the remaining 4 for testing.  The best performing channel is informed as well.  It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}

>>>>

This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained. Section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.  Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.  Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.

@@

The P300~\citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.  It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and it is consistent across different subjects. It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\citep{Hu2010}.  This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}. Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.   Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.

@@

The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
The processing pipeline starts by applying a notch filter to the raw signal, a 
$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.

@@

The aim is to identify the selected letter from the matrix. Previously, during training phase, descriptors from \textit{hit} labeled segments are extracted.  These descriptors are the P300 templates which are grouped in a template set called $ T $.  Segments corresponding to rows are labeled 1-6, whereas those corresponding to columns are labeled 7-12.  The process has the following steps:

\begin{enumerate}
\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively. The parameter $k_a$ is an input of the algorithm, and corresponds to the number of repetitions.
\item Compute  
\begin{equation}
x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
\label{averaging}
\end{equation}

@@

\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ in accordance to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.

\item Match to the Template $T$ by computing

@@

where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.  By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.

@@

Figure~\ref{fig:classification} shows a schema of this process. 
\begin{figure}
\centering
\includegraphics[width=15cm]{classificationgraph2.pdf}
\caption{Single trial segments are averaged for the 6 rows and 6 columns.}
\label{fig:classification}
\end{figure}

@@

To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, is used. Additionally, an own dataset with  the same experimental conditions is generated. Both of them are utilized to perform an offline BCI Simulation to decode the spelled words from the provided signals.

The algorithm is implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).

@@

\subsubsection{P300 ALS Public Dataset} \label{ALSDataset}

@@

The recorded dataset was sampled at 256 Hz and it consisted of scalp EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.   The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\citep{Schalk2004}.

@@

We replicated the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).  The experimental conditions were the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.

Participants were recruited voluntarily and the experiment were conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.  No monetary compensation was handed out and all participants agreed and signed a written informed consent.  All healthy subjects had normal or corrected-to normal vision and no history of neurological disorders. The experiment was performed with 20 subjects, 7 males, 4 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.

Gel-based active electrodes (g.LADYbird, g.Tec, Austria) were used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.  Reference was set to the right ear lobe and ground was preset as the AFz position.   Sampling frequency is slightly different, and was set to 250 Hz, which is the closest possible to the one used with the other dataset.

%EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems. The reference positions were placed at the right ear. The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed). Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz). Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz. The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system. Table~\ref{tab:results} summarizes the acquisition settings for each system.

@@

% Antes estaba esto aca

% The P300 event latency and amplitude vary greatly between subjects,

% lo saque porque confunde, ya que despues de todo, dejamos todo fijo, e ignoramos completamente el hecho de que varie.

Parameters are selected according to the experimental protocol. First of all, the patch size in pixels is $S_P = 3s.4 \times 3s.4$. The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.  Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event. With this purpose in consideration, the $s$ value election is essential.
%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
\begin{equation}
s_x = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
\label{eq:mapping2}
\end{equation}

@@

\noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch. The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.  This also provides a resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs \cdot \gamma}$ seconds in the horizontal direction. Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).   In this way the whole transient event should be captured.

% esta figura no está comentada en el texto.
\begin{figure}[H]
\centering
\includegraphics[width=10cm]{patchgeometry.pdf}
\caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
\label{fig:sampledescriptor2}
\end{figure}

@@

Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.  From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.  The best performing channel is informed as well.

DRAFT: Traditionally P300 can be seen in this channel and this other channel.  However, Wolpaw showed that the response may also arise in occipital channels as well.  The authors of the original paper concluded that Cz was the best performing channel.  However, in our approach

It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University. The authors would like to thank Dr. Juliana Gambini for their insights and help with the detailed description of the histogram of gradients.}
